Fourmile Calibrations and Output

### 2025-09-02 fourmile_output_dec2024.Rmd
cleaning up fig4 - moved to manuscriptFig Rmd

write_csv(nseCats, '/Users/megansears/Documents/Repos/fourmile/38HRU_dec2024_updates/nseCats.csv')

@@

@@

write_csv(flow_plot,
          '/Users/megansears/Documents/Repos/fourmile/38HRU_dec2024_updates/all_kge.csv')

@@

both_kge_plot <- flow_plot_updated %>%
  ggplot(., aes(x = year, 
              y = year_excluded, 
              fill = kge_revise_bins)) +
  geom_tile(color = "white") +
  facet_wrap(~source, labeller = labeller(source = new_facet_titles)) +
  #paletteer::scale_fill_paletteer_d("PNWColors::Anemone") +
  # paletteer::scale_fill_paletteer_d(
  # "PNWColors::Anemone", 
  # guide = guide_legend(nrow = 1),
  # direction = -1) +
   scale_fill_manual(values = c(
    "< 0.0" = "#DCBE9BFF", 
    "0.0-0.2" = "#EFDDCFFF",
    "0.2-0.4" = "#F1F4EEFF",
    "0.4-0.6" = "#72E1E1FF",
    "0.6-0.8" = "#11C2B5FF",
    "> 0.8" = "#009474FF"),
    guide = guide_legend(nrow = 1)) +
  geom_text(aes(label = sprintf("%.2f", 
                                kge)), 
            color = "black", 
            size = 3) +
  theme_bw(base_size=18) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1)) +
  labs(y = "Year excluded",
       x='Year',
       fill = 'KGE') +
  theme(
    axis.text.x = element_text(angle = 45, 
                               hjust = 1),
    legend.position = 'bottom',
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.key.width = unit(.6, "cm"),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    legend.margin = margin(t = -8, b = -4, unit = "pt"),
  )

@@

ggsave('/Users/megansears/Documents/Repos/fourmile/figures/kge_heatmap_both_basic.png',
       dpi=600,
       width = 10,
       height = 6)

@@

write_csv(pbiasCats, '/Users/megansears/Documents/Repos/fourmile/38HRU_dec2024_updates/pbiasCats.csv')

both_pbias_plot <- pbiasCats %>%
  ggplot(aes(x = year, 
             y = year_excluded, 
             fill = pbiasCat)) +
  geom_tile(color = "white") +
  facet_wrap(~source, labeller = labeller(source = new_facet_titles)) +
   scale_fill_manual(values = c(
    "Unsatisfactory" = "#F1F4EEFF", 
    "Satisfactory" = "#EFDDCFFF",
    "Good" = "#72E1E1FF",
    "Very good" = "#009474FF"
  )) +
  geom_text(aes(label = sprintf("%.2f", pbias)), 
            color = "black", 
            size = 2.75) +
  theme_bw(base_size=18) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = 'bottom',
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.key.width = unit(1, "cm"),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    legend.margin = margin(t = -8, b = -4, unit = "pt")
  ) +
  labs(y = "Year excluded",
       x = 'Year',
       fill = '% bias')

@@

ggsave('/Users/megansears/Documents/Repos/fourmile/figures/pbias_heatmap_both_cats.png',
       dpi=600,
       width = 10,
       height = 6)

>>>>

#write_csv(nseCats, '/Users/megansears/Documents/Repos/fourmile/38HRU_dec2024_updates/nseCats.csv')

@@

both_nse_plot <- nseCats %>%
  ggplot(aes(x = year, 
             y = year_excluded, 
             fill = nseCat)) +
  geom_tile(color = "white") +
  facet_wrap(~source, labeller = labeller(source = new_facet_titles)) +
  scale_fill_manual(values = c(
    "Unsatisfactory" = "#F1F4EEFF", 
    "Satisfactory"  = "#EFDDCFFF",
    "Good"          = "#72E1E1FF",
    "Very good"     = "#009474FF"
  )) +
  geom_text(aes(label = sprintf("%.2f", nse), color = nseCat), 
            size = 2.75) +
  scale_color_manual(values = c(
    "Unsatisfactory" = "black",
    "Satisfactory"   = "black",
    "Good"           = "black",
    "Very good"      = "white"
  ), guide = "none") +
  theme_bw(base_size=18) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = 'bottom',
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.key.width = unit(1, "cm"),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    legend.margin = margin(t = -8, b = -4, unit = "pt")
  ) +
  labs(y = "Year excluded",
       x = 'Year',
       fill = 'NSE')

ggsave('/Users/megansears/Documents/Repos/fourmile/figures/nse_heatmap_both_cats_whiteft.png',
       dpi=600,
       width = 10,
       height = 6)

@@

# write_csv(flow_plot,
#           '/Users/megansears/Documents/Repos/fourmile/38HRU_dec2024_updates/all_kge.csv')

@@

both_kge_plot <- flow_plot_updated %>%
  ggplot(., aes(x = year, 
              y = year_excluded, 
              fill = kge_revise_bins)) +
  geom_tile(color = "white") +
  facet_wrap(~source, labeller = labeller(source = new_facet_titles)) +
   scale_fill_manual(values = c(
    "< 0.0" = "#DCBE9BFF", 
    "0.0-0.2" = "#EFDDCFFF",
    "0.2-0.4" = "#F1F4EEFF",
    "0.4-0.6" = "#72E1E1FF",
    "0.6-0.8" = "#11C2B5FF",
    "> 0.8" = "#009474FF"),
    guide = guide_legend(nrow = 1)) +
  # geom_text(aes(label = sprintf("%.2f", 
  #                               kge)), 
  #           color = "black", 
  #           size = 3) +
    geom_text(aes(label = sprintf("%.2f", kge), 
                color = kge_revise_bins), 
            size = 3) +
  scale_color_manual(values = c(
    "< 0.0"   = "black",
    "0.0-0.2" = "black",
    "0.2-0.4" = "black",
    "0.4-0.6" = "black",
    "0.6-0.8" = "black",
    "> 0.8"   = "white"
  ), guide = "none") +
  theme_bw(base_size=18) +
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1)) +
  labs(y = "Year excluded",
       x='Year',
       fill = 'KGE') +
  theme(
    axis.text.x = element_text(angle = 45, 
                               hjust = 1),
    legend.position = 'bottom',
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.key.width = unit(.6, "cm"),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    legend.margin = margin(t = -8, b = -4, unit = "pt"),
  )

@@

ggsave('/Users/megansears/Documents/Repos/fourmile/figures/kge_heatmap_whiteft.png',
       dpi=800,
       width = 10,
       height = 6)

@@

#write_csv(pbiasCats, '/Users/megansears/Documents/Repos/fourmile/38HRU_dec2024_updates/pbiasCats.csv')

both_pbias_plot <- pbiasCats %>%
  ggplot(aes(x = year, 
             y = year_excluded, 
             fill = pbiasCat)) +
  geom_tile(color = "white") +
  facet_wrap(~source, labeller = labeller(source = new_facet_titles)) +
   scale_fill_manual(values = c(
    "Unsatisfactory" = "#F1F4EEFF", 
    "Satisfactory" = "#EFDDCFFF",
    "Good" = "#72E1E1FF",
    "Very good" = "#009474FF"
  ),
  #guide = guide_legend(ncol = 1)
  ) +
  # geom_text(aes(label = sprintf("%.2f", pbias)), 
  #           color = "black", 
  #           size = 2.75) +
  geom_text(aes(label = sprintf("%0.1f", pbias),   # one decimal place
                color = pbiasCat), 
            size = 2.75) +
  scale_color_manual(values = c(
    "Unsatisfactory" = "black",
    "Satisfactory"   = "black",
    "Good"           = "black",
    "Very good"      = "white"
  ), guide = "none") + 
  theme_bw(base_size=18) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = 'bottom',
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.key.width = unit(1, "cm"),
    panel.grid = element_blank(),
    panel.background = element_rect(fill = "white"),
    strip.background = element_blank(),
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10),
    legend.margin = margin(t = -8, b = -4, unit = "pt")
  ) +
  labs(y = "Year excluded",
       x = 'Year',
       fill = '% bias')

@@

ggsave('/Users/megansears/Documents/Repos/fourmile/figures/pbias_heatmap_whiteft.png',
       dpi=800,
       width = 10,
       height = 6)

plot_grid(both_kge_plot,
          both_pbias_plot,
          both_nse_plot,
          ncol=1,
          labels = c("a) KGE", 
                     "b) Bias", 
                     "c) NSE"),
          label_size = 18,
          label_fontface = "bold",
          label_x = 0.02,
          label_y = 0.98
)

ggsave('/Users/megansears/Documents/Repos/fourmile/figures/heatmap_metrics.png',
       dpi=800,
       width = 10,
       height = 14)

########

# Top plot
top_plot <- both_kge_plot +
  theme(
    axis.title.x = element_blank(),
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none",
    plot.margin = margin(5, 5, 5, 40)   # add space on the left
  )

# Middle plot
middle_plot <- both_pbias_plot +
  theme(
    strip.text = element_blank(),
    axis.title.x = element_blank(),
    axis.text.x  = element_blank(),
    axis.ticks.x = element_blank(),
    legend.position = "none",
    plot.margin = margin(5, 5, 5, 40)   # add space on the left
  )

# Bottom plot
bottom_plot <- both_nse_plot +
  theme(
    strip.text = element_blank(),
    legend.position = "none",
    plot.margin = margin(5, 5, 5, 40)   # add space on the left
  )

# Combine plots
final_grid <- plot_grid(
  top_plot,
  middle_plot,
  bottom_plot,
  ncol = 1,
  labels = c("a) KGE", "b) Bias", "c) NSE"),
  label_size = 16,
  label_fontface = "bold",
  label_x = -0.03,
  label_y = 1.04,
  rel_heights = c(1.1, 1, 1.2),
  align = "v",
  axis = "l"
)

final_grid_add <- final_grid +
  theme(plot.margin = margin(20, 5, 5, 5))

ggsave('/Users/megansears/Documents/Repos/fourmile/figures/heatmap_metrics.pdf',
       dpi=800,
       width = 10,
       height = 14,
       bg='white')
